---
title: "Downstream analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

```{r}
library(Seurat)
library(ggplot2)
```

## Load data

```{r}
#count_file='/Users/xiaz/1My_Work/2020_single_cell_subpopulation/Faldmen_T_cell_Umap/Feldman_T_cell_scaled_data_exp.csv'
#raw.data=read.csv(count_file,stringsAsFactors=F)
#dim(raw.data)
#raw.data[1:5,1:5]
rm(list=ls())
load('./Feldman_T_cell_Seurat_object.RData')
pbmc <- CreateSeuratObject(counts = pbmc@assays[["RNA"]]@counts, meta.data = pbmc@meta.data)
pbmc <- NormalizeData(object=pbmc, normalization.method="LogNormalize")

#data <- NormalizeData(object=data, normalization.method="LogNormalize")
#data <- NormalizeData(object=data, normalization.method="RC", scale.factor=1000000)

#if(counts) data <- NormalizeData(object=data, normalization.method='LogNormalize', scale.factor=10000)
#data <- FindVariableFeatures(object=data, selection.method='mean.var.plot', mean.cutoff=c(0.0125, 3), dispersion.cutoff=c(0.5, Inf))
#data <- FindVariableFeatures(object=data, selection.method='vst', nfeatures=2000)
#var_features_genes=data@assays$RNA@var.features

pbmc <- FindVariableFeatures(object=pbmc, selection.method='vst', nfeatures=2000)
pbmc <- ScaleData(object=pbmc)
```


```{r}
pbmc <- RunPCA(object=pbmc, features=VariableFeatures(pbmc)[1000:1300])
ElbowPlot(pbmc)
# write.csv(t(pbmc@reductions[["pca"]]@cell.embeddings), './exp_data_pca50_1000_1300.csv'4

pbmc <- FindNeighbors(object=pbmc, dims=1:10, k.param=20)
pbmc <- FindClusters(object=pbmc, resolution=1.0, algorithm = 1)
#data <- RunTSNE(object=data, dims=1:10)
pbmc <- RunUMAP(object=pbmc, dims=1:10)

#DimPlot(object=pbmc, reduction='umap', label=T,group.by='responsive_status', label.size=10)
# DimPlot(object=pbmc, reduction='umap', label=F, group.by='responsive_status')
DimPlot(object=pbmc, reduction='umap', label=T, label.size=10)
ggsave('./results/1000-1300-umap.pdf')
# selected_genes = VariableFeatures(pbmc)[c(999,  1100,  1279) + 1]
# FeaturePlot(object=pbmc, reduction="umap", features=selected_genes)
# VlnPlot(object=pbmc, features=selected_genes, group.by=true_groups_simulation_name, pt.size=0.01)

# pbmc <- FindClusters(object=pbmc, resolution=1.0)
# DimPlot(object=pbmc, reduction='umap', label=T, label.size=10)
# DimPlot(object=pbmc, reduction='pca', label=F, label.size=10)
# ## DEG comparing
# DEG_result <- FindMarkers(object=pbmc, ident.1="Responder", ident.2="Non-responder", group.by='responsive_status', logfc.threshold=log(1.25))

# embeddings = pbmc@reductions[["umap"]]@cell.embeddings
# write.csv(embeddings, './embedding-umap-middle2.csv')

# pbmc <- RunUMAP(object=pbmc, features = VariableFeatures(pbmc)[700:1000])
# DimPlot(object=pbmc, reduction='umap', label=T)
# pbmc <- RunUMAP(object=pbmc, dims=1:50)
# DimPlot(object=pbmc, reduction='umap', label=T)
```

# Generated 2-class labels
```{r}
simulation_name = 'simulation'
# table(pbmc$RNA_snn_res.1.6)

mix_rate1 = 0.1
mix_rate2 = 0.1
#phenotype 1
# table(pbmc$RNA_snn_res.1)
selected_cells_phenotype1 = c(colnames(pbmc)[pbmc$RNA_snn_res.1==3],
                              colnames(pbmc)[pbmc$RNA_snn_res.1==14])
                              # colnames(pbmc)[pbmc$RNA_snn_res.1==5])
                              #colnames(pbmc)[pbmc$RNA_snn_res.1.2==3])
# phenotype 2
# table(pbmc$RNA_snn_res.1)
selected_cells_phenotype2 = c(colnames(pbmc)[pbmc$RNA_snn_res.1==5])
                              # colnames(pbmc)[pbmc$RNA_snn_res.1==10])
                              # colnames(pbmc)[pbmc$RNA_snn_res.0.8==13])
                              #colnames(pbmc)[pbmc$RNA_snn_res.1.2==12])

#phenotype 1
# table(pbmc$RNA_snn_res.1)
# selected_cells_phenotype1 = c(colnames(pbmc)[pbmc$RNA_snn_res.1==3])
#                               # colnames(pbmc)[pbmc$RNA_snn_res.1==3])
#                               # colnames(pbmc)[pbmc$RNA_snn_res.1==5])
#                               #colnames(pbmc)[pbmc$RNA_snn_res.1.2==3])
# #phenotype2
# table(pbmc$RNA_snn_res.1)
# selected_cells_phenotype2 = c(colnames(pbmc)[pbmc$RNA_snn_res.1.5==12])


selected_phenotypic_cells = rep('Group1', ncol(pbmc))
names(selected_phenotypic_cells) = colnames(pbmc)
selected_phenotypic_cells[selected_cells_phenotype1] = "Group2"
selected_phenotypic_cells[selected_cells_phenotype2] = "Group3"

true_groups_simulation_name = sprintf("true_groups_%s",simulation_name)
pbmc@meta.data[[true_groups_simulation_name]] <- selected_phenotypic_cells

B = DimPlot(pbmc,group.by=true_groups_simulation_name,cols=c("gray","#3283FE","red"),order = c("Group3","Group2"),label = T)
ggsave('./results/group_in_300.pdf')
print(B)

# write.csv(selected_phenotypic_cells, file=sprintf('/Users/xiaz/1My_Work/2020_single_cell_subpopulation/Feldman_simulation/Feldman_simulated_data/Feldman_data_true_groups_%s.csv',simulation_name),quote=F)
##labels the cells with the phenotypes

#labels_res <- c("1")
#labels_nonres <- c("2")
labels = rep(-1,dim(pbmc)[2])
names(labels) = colnames(pbmc)
simulated_true_groups = pbmc@meta.data[[true_groups_simulation_name]]
table(simulated_true_groups)


num_group1 = sum(simulated_true_groups=="Group1")
tmp = floor(num_group1/2)
tmp_ids = sample.int(num_group1, tmp)
labels[simulated_true_groups=="Group1"] = 'class_1'
labels[simulated_true_groups=="Group1"][tmp_ids] = 'class_2'
labels[simulated_true_groups=="Group1"] = labels[simulated_true_groups=="Group1"][sample.int(num_group1, num_group1)]


num_group2 = sum(simulated_true_groups=="Group2")#NR
tmp = floor(num_group2* mix_rate1)
tmp_ids = sample.int(num_group2, tmp)
labels[simulated_true_groups=="Group2"] = 'class_1'
labels[simulated_true_groups=="Group2"][tmp_ids] = 'class_2'

num_group3 = sum(simulated_true_groups=="Group3")
tmp = floor(num_group3 * mix_rate2)
# tmp = floor(num_group3/2)
tmp_ids = sample.int(num_group3, tmp)
labels[simulated_true_groups=="Group3"] = 'class_2'
labels[simulated_true_groups=="Group3"][tmp_ids] = 'class_1'

table(labels, simulated_true_groups)


pbmc <- AddMetaData(object=pbmc, metadata=labels, col.name=sprintf("cell_phenotype_labels_%s",simulation_name))


A = DimPlot(pbmc,group.by=sprintf("cell_phenotype_labels_%s",simulation_name),cols=c("#3283FE","red"))
ggsave('./results/class_label_in_300.pdf')
print(A)
# write.csv(labels, file=sprintf('/Users/xiaz/1My_Work/2020_single_cell_subpopulation/Feldman_simulation/Feldman_simulated_data/Feldman_%s_phenotype_labels.csv',simulation_name),quote=F)

# save(pbmc, file=sprintf('/Users/xiaz/1My_Work/2020_single_cell_subpopulation/Feldman_simulation/Feldman_simulated_data/Feldman_%s_seurat_obj_with_phenotype_labels.RData',simulation_name))
#a=pbmc@reductions$pca@cell.embeddings

embeddings = pbmc@reductions[["umap"]]@cell.embeddings
labels = as.matrix(labels)

# selected_genes = VariableFeatures(pbmc)[1:1000]
# exp = pbmc@assays$RNA@scale.data
# exp_for_selected = exp[selected_genes, ]
# exp_for_selected = as.matrix(exp_for_selected)

# write.csv(exp_for_selected, './exp_data_1000.csv')
write.csv(labels, './label_info_2s.csv')
# write.csv(embeddings, './embedding.csv')

# DimPlot(pbmc,group.by=true_groups_simulation_name,cols=c("gray","#3283FE","red"),order = c("Group3","Group2"),label = T, reduction = 'umap')

```
# umap for 2000 variable genes
```{r}
pbmc_ <- pbmc
pbmc_ <- RunPCA(object=pbmc_, features=VariableFeatures(pbmc_), npcs=20)

pbmc_ <- FindNeighbors(object=pbmc_, dims=1:20, k.param=20)
pbmc_ <- FindClusters(object=pbmc_, resolution=1.0)
#data <- RunTSNE(object=data, dims=1:10)
pbmc_ <- RunUMAP(object=pbmc_, dims=1:20)

# write.csv(t(pbmc@reductions[["pca"]]@cell.embeddings), './exp_data_pca50.csv')

ElbowPlot(pbmc_, 20)
C = DimPlot(pbmc_,group.by=true_groups_simulation_name,cols=c("gray","#3283FE","red"),order = c("Group3","Group2"),label = T)
print(C)
ggsave('./results/groups_in_top2000.pdf')
DimPlot(pbmc_,group.by=sprintf("cell_phenotype_labels_%s",simulation_name),cols=c("#3283FE","red"),order = c("class_2","class_1"),label = T)
ggsave('./results/class_label_in_top2000.pdf')
DimPlot(object=pbmc_, reduction='umap', label=T, label.size=10)
ggsave('./results/clusters_of_top2000.pdf')
pbmc@reductions[['umap-2000']] <- pbmc_@reductions[['umap']]

```

# Add the prediction result - bi-classification
```{r}
library(plyr)
library(VennDiagram)
library(grid)
save_path = './results/'

input_genes_id = 1:2000
expr_id = '2s_1e-5_100epochs'

# expr_id = "pca10-umap_radom_rate_0.2_radom_rate_0.3_radom_rate_0.4_radom_rate_0.5_radom_rate_0.6_radom_rate_0.7"
prediction_result_file = sprintf('../../results/Feldman_T_cell_for_select_genes/predicted_labels_%s.csv', expr_id)
scores_result_file = sprintf('../../results/Feldman_T_cell_for_select_genes/high_dimens_result_%s.csv', expr_id)
predict_lables=read.csv(prediction_result_file, stringsAsFactors=F)
scores = read.csv(scores_result_file, header = F)
row.names(scores) = colnames(pbmc)
colnames(scores) = c('real', 'predictd', 'confidence.score')
pbmc <- AddMetaData(object=pbmc, metadata=scores)


dim(predict_lables)
table(predict_lables[, 2])
rownames(predict_lables)=predict_lables[, 1]
all(predict_lables[, 1]==colnames(pbmc))
predict_lables_meta = predict_lables[, 2]
# predict_lables_meta = mapvalues(predict_lables_meta, c("Responder","Non-responder", "Rejected" ), c(2, 1, 0))
#names(predict_lables) <- colnames(pbmc)
#read 
pbmc <- AddMetaData(object=pbmc, metadata=predict_lables_meta, col.name="predicted_lables")
D = DimPlot(pbmc,group.by="predicted_lables",cols=c("gray","#3283FE","red"),order = c("class_2","class_1"),label = T)
ggsave(paste(save_path,  "predicted_labels.pdf", sep=''))
print(D)
C = FeaturePlot(pbmc, features = 'confidence.score')
ggsave('./results/confidence.score.pdf')
print(C)

write.csv(pbmc$predicted_lables, file = './results/pencil_predicted_labels.csv')

num_g1 = sum(pbmc@meta.data[[true_groups_simulation_name]] == 'Group1')
num_g2 = sum(pbmc@meta.data[[true_groups_simulation_name]] == 'Group2')
num_g3 = sum(pbmc@meta.data[[true_groups_simulation_name]] == 'Group3')

num_rej_in_g1 = sum((pbmc@meta.data[[true_groups_simulation_name]] == 'Group1') & (pbmc@meta.data[["predicted_lables"]] == 'Rejected'))
num_rej_in_g2 = sum((pbmc@meta.data[[true_groups_simulation_name]] == 'Group2') & (pbmc@meta.data[["predicted_lables"]] == 'Rejected'))
num_rej_in_g3 = sum((pbmc@meta.data[[true_groups_simulation_name]] == 'Group3') & (pbmc@meta.data[["predicted_lables"]] == 'Rejected'))

overlap_of_rej = data.frame(
  Group1.Background = c(num_rej_in_g1, num_g1 - num_rej_in_g1), 
  Group2 = c(num_rej_in_g2, num_g2 - num_rej_in_g2), 
  Group3 = c(num_rej_in_g3, num_g3 - num_rej_in_g3), 
  row.names = c("Rejected", "Not_rejected")
)

df = data.frame(
  frequency = as.vector(t(as.matrix(overlap_of_rej))),
  type = rep(colnames(overlap_of_rej), 2),
  rejected = c(rep(c("Y"), 3), rep(c("N"), 3))
)
df$type = factor(df$type, levels=c('Group1.Background','Group2','Group3'))
df$rejected = factor(df$rejected, levels=c('Y','N'))
ggplot(data = df, mapping = aes(x = type, y = frequency, fill = rejected)) + geom_bar(stat = 'identity', position = 'stack', width = 0.5)
ggsave(paste(save_path,  "overlap.pdf", sep=''))
ggplot(data = df, mapping = aes(x = type, y = frequency, fill = rejected)) + geom_bar(stat = 'identity', position = 'fill', width = 0.5)
ggsave(paste(save_path,  "overlap_scale.pdf", sep=''))

cell_ids = colnames(pbmc)

ggplot()
venn.plot1 <- venn.diagram(
    x = list(
        "Eriched in phenotype I" = cell_ids[pbmc@meta.data[[true_groups_simulation_name]] == 'Group2'],
        "Pencil predicted subpopulation 1" = cell_ids[pbmc@meta.data[["predicted_lables"]] == 'class_1']
    ),
    filename = NULL,
    col = "transparent",
    fill = c("skyblue", "orange"),
    cex = 2,
    cat.cex = 1,
    rotation.degree = 0,
    main = "",
    main.cex = 2,
    sub.cex = 1,
    alpha = 0.50,
    # cat.pos=c(9,12)
)
grid.draw(venn.plot1)
ggsave(paste(save_path,  "venn_g1.pdf", sep=''))

ggplot()
venn.plot2 <- venn.diagram(
    x = list(
        "Eriched in phenotype II" = cell_ids[pbmc@meta.data[[true_groups_simulation_name]] == 'Group3'],
        "Pencil predicted subpopulation 2" = cell_ids[pbmc@meta.data[["predicted_lables"]] == 'class_2']
    ),
    filename = NULL,
    col = "transparent",
    fill = c("skyblue", "orange"),
    cex = 2,
    cat.cex = 1,
    rotation.degree = 0,
    main = "",
    main.cex = 2,
    sub.cex = 1,
    alpha = 0.50,
    # cat.pos=c(9,12)
)
grid.draw(venn.plot2)
ggsave(paste(save_path,  "venn_g2.pdf", sep=''))

## DEG comparing
# DEG_result_confident <- FindMarkers(object=pbmc, ident.1=2, ident.2=1, group.by='predict_lables', logfc.threshold=log(1.25))

for (each in list.files('./')){
  if (substring(each, nchar(each)-2) == 'log'){
    file.remove(each)
  }
}


library("cowplot")
plot_grid(A, B, C, D, labels = c("A", "B", "C", "D"), ncol = 2, nrow = 2)
ggsave('./results/pencil_results.pdf', width = 11, height = 8)


plot_grid(venn.plot1, venn.plot2, labels = c("E","F"), ncol = 2, nrow = 1)
ggsave('./results/venn.pdf', width = 15, height = 4)


gene_weights_file = sprintf('../../results/Feldman_T_cell_for_select_genes/selected_weight_%s.csv', expr_id)
gene_weights = read.csv(gene_weights_file, header = F)
colnames(gene_weights) = 'weight'
gene_weights$gene = row.names(gene_weights)
ggplot(data=gene_weights, mapping=aes(x=gene, y=weight))+geom_bar(stat="identity")+scale_x_discrete(limits=1:2000, breaks=seq(0,2000, 200))
ggsave('./results/weights_of_genes.pdf')

seleted_ids = order(-abs(gene_weights$weight))[1:12]
selected_genes = VariableFeatures(pbmc)[input_genes_id][seleted_ids]
FeaturePlot(object=pbmc, reduction="umap", features=selected_genes)
ggsave('./results/genesplot.pdf', width = 20, height = 20)
VlnPlot(object=pbmc, features=selected_genes, group.by=true_groups_simulation_name, pt.size=0.01)

FeaturePlot(pbmc, features = 'confidence.score', cells = colnames(pbmc)[pbmc$confidence.score > 0.5])

gene_weights$method = rep('pencil', nrow(gene_weights))
gene_weights$weight.abs = abs(gene_weights$weight)
ggplot(data = gene_weights, aes(x=gene, y=method, fill=weight)) + geom_tile()+scale_x_discrete(limits=1:2000, breaks=seq(0,2000, 200)) + scale_fill_gradient2(low='blue', mid='black', high='red')
ggsave('./results/weights_of_genes_heatmap.pdf')
```

# analysis for selected genes
```{r}
seleted_ids = abs(gene_weights$weight) > 0.5
selected_genes = VariableFeatures(pbmc)[input_genes_id][seleted_ids]

pbmc_selected <- pbmc[selected_genes, ]
pbmc_selected  <- ScaleData(object=pbmc_selected)
pbmc_selected <- RunPCA(object=pbmc_selected, features=selected_genes, npcs=20)

pbmc_selected <- FindNeighbors(object=pbmc_selected, dims=1:10, k.param=20)
pbmc_selected <- FindClusters(object=pbmc_selected, resolution=0.8)
# data <- RunTSNE(object=data, dims=1:10)
pbmc_selected <- RunUMAP(object=pbmc_selected, dims=1:10)

# pbmc_selected <- RunUMAP(object=pbmc_selected, features=rownames(pbmc_selected))
# write.csv(t(pbmc@reductions[["pca"]]@cell.embeddings), './exp_data_pca50.csv')

ElbowPlot(pbmc_selected, 30)
# DimPlot(pbmc_selected, group.by="predicted_lables", cols=c("gray","#3283FE","red"),order = c("class_2","class_1"),label = T)

DimPlot(pbmc_selected, group.by="predicted_lables", cols=c("gray", "#3283FE","red"),order = c("class_2","class_1"),label = T)
ggsave('./results/pencil_predict_for_selected_genes-umap.pdf')
DimPlot(pbmc_selected,group.by="true_groups_simulation",cols=c("gray","#3283FE","red"),order = c("Group3","Group2"),label = T)
ggsave('./results/true_group_for_selected_genes-umap.pdf')
# DimPlot(object=pbmc_selected, reduction='umap', label=T, label.size=10)
FeaturePlot(pbmc_selected, features = 'confidence.score')
ggsave('./results/confidence_score_for_selected_genes.pdf')


```


# analysis for selected genes and cells
```{r}
seleted_ids = abs(gene_weights$weight) > 0.1
selected_genes = VariableFeatures(pbmc)[input_genes_id][seleted_ids]

pbmc_selected <- pbmc[selected_genes, pbmc$predicted_lables!='Rejected']
pbmc_selected <- RunPCA(object=pbmc_selected, features=selected_genes, npcs=20)

pbmc_selected <- FindNeighbors(object=pbmc_selected, dims=1:20, k.param=20)
pbmc_selected <- FindClusters(object=pbmc_selected, resolution=0.8)
# data <- RunTSNE(object=data, dims=1:10)
pbmc_selected <- RunUMAP(object=pbmc_selected, dims=1:20)

# pbmc_selected <- RunUMAP(object=pbmc_selected, features=rownames(pbmc_selected))
# write.csv(t(pbmc@reductions[["pca"]]@cell.embeddings), './exp_data_pca50.csv')

ElbowPlot(pbmc_selected, 30)
# DimPlot(pbmc_selected, group.by="predicted_lables", cols=c("gray","#3283FE","red"),order = c("class_2","class_1"),label = T)

DimPlot(pbmc_selected, group.by="predicted_lables", cols=c("#3283FE","red"),order = c("class_2","class_1"),label = T)
DimPlot(pbmc_selected,group.by="true_groups_simulation",cols=c("gray","#3283FE","red"),order = c("Group3","Group2"),label = T)
ggsave('./results/umap_selected_genes_and_selected_cells.pdf')
DimPlot(object=pbmc_selected, reduction='umap', label=T, label.size=10)
FeaturePlot(pbmc_selected, features = 'confidence.score')

```


# analysis for selected cells
```{r}
selected_genes = VariableFeatures(pbmc)[input_genes_id]
pbmc_selected <- pbmc[selected_genes, pbmc$predicted_lables!='Rejected']
pbmc_selected <- RunPCA(object=pbmc_selected, features=selected_genes, npcs=20)

pbmc_selected <- FindNeighbors(object=pbmc_selected, dims=1:20, k.param=20)
pbmc_selected <- FindClusters(object=pbmc_selected, resolution=0.8)
# data <- RunTSNE(object=data, dims=1:10)
pbmc_selected <- RunUMAP(object=pbmc_selected, dims=1:20)

# pbmc_selected <- RunUMAP(object=pbmc_selected, features=rownames(pbmc_selected))
# write.csv(t(pbmc@reductions[["pca"]]@cell.embeddings), './exp_data_pca50.csv')

ElbowPlot(pbmc_selected, 30)
# DimPlot(pbmc_selected, group.by="predicted_lables", cols=c("gray","#3283FE","red"),order = c("class_2","class_1"),label = T)

DimPlot(pbmc_selected, group.by="predicted_lables", cols=c("#3283FE","red"),order = c("class_2","class_1"),label = T)
DimPlot(pbmc_selected,group.by="true_groups_simulation",cols=c("gray","#3283FE","red"),order = c("Group3","Group2"),label = T)
ggsave('./results/umap_selected_selected_cells.pdf')
DimPlot(object=pbmc_selected, reduction='umap', label=T, label.size=10)
FeaturePlot(pbmc_selected, features = 'confidence.score')

```
# Pencil DEG genes
```{r}

# cells.1 = Cells(pbmc)[pbmc$predicted_lables=='class_1']
# cells.2 = Cells(pbmc)[pbmc$predicted_lables=='class_2']
# 
# deg <- FindMarkers(object=pbmc, ident.1=cells.1, ident.2=cells.2, group.by='predicted_lables', logfc.threshold = log(1.25))

deg <- FindMarkers(object=pbmc, ident.1='class_1', ident.2='class_2', group.by='predicted_lables', logfc.threshold = log(1.25))

features <- row.names(deg)[deg$p_val<0.05]
features <- unique(features)
# pbmc <- pbmc[features, ]
#
# pbmc <- NormalizeData(object=pbmc, normalization.method="LogNormalize")
# # pbmc <- FindVariableFeatures(object=pbmc, selection.method='dispersion', nfeatures=2000)
pbmc <- ScaleData(object=pbmc, features = features)
pbmc <- RunPCA(object=pbmc, features=features)
ElbowPlot(pbmc)
# pbmc <- RunPCA(object=pbmc, npcs=30)
pbmc <- FindNeighbors(object=pbmc, dims=1:20, k.param=20)
pbmc <- FindClusters(object=pbmc, resolution=0.8)
pbmc <- RunUMAP(object=pbmc, dims=1:20)

DimPlot(object=pbmc, reduction='umap', group.by="predicted_lables", cols=c("gray", "#3283FE","red"),order = c("class_2","class_1"), label = T)

ggsave('./results/umap_deg_from_pencil_predict_lables.pdf')
```
#raw sample condition deg
```{r}
deg <- FindMarkers(object=pbmc, ident.1='class_1', ident.2='class_2', group.by='cell_phenotype_labels_simulation', logfc.threshold = 0.1)

features <- row.names(deg)[deg$p_val<0.05]
features <- unique(features)
# pbmc <- pbmc[features, ]
#
# pbmc <- NormalizeData(object=pbmc, normalization.method="LogNormalize")
# # pbmc <- FindVariableFeatures(object=pbmc, selection.method='dispersion', nfeatures=2000)
pbmc <- ScaleData(object=pbmc, features = features)
pbmc <- RunPCA(object=pbmc, features=features)
ElbowPlot(pbmc)
# pbmc <- RunPCA(object=pbmc, npcs=30)
pbmc <- FindNeighbors(object=pbmc, dims=1:20, k.param=20)
pbmc <- FindClusters(object=pbmc, resolution=0.8)
pbmc <- RunUMAP(object=pbmc, dims=1:20)

DimPlot(object=pbmc, reduction='umap', group.by="predicted_lables", cols=c("gray", "#3283FE","red"),order = c("class_2","class_1"), label = T)

DimPlot(object=pbmc, reduction='umap', group.by="cell_phenotype_labels_simulation", cols=c("#3283FE","red"), order = c("class_2","class_1"), label = T)
ggsave('./results/umap_deg_from_simulated_true_labels.pdf')
```


# analysis for selected cells
##funciton for calculate overlap
```{r}
library(VennDiagram)
library(grid)
library(cowplot)
overlap_analysis <- function(pbmc, pos_check, select_check, method='pencil'){
    cell_ids = colnames(pbmc)
    
    venn.plot <- venn.diagram(
        x = list(
            "ground truth" = cell_ids[pos_check],
            "predicted subpopulation" = cell_ids[select_check]
        ),
        filename = NULL,
        col = "transparent",
        fill = c("skyblue", "orange"),
        cex = 2,
        cat.cex = 1,
        rotation.degree = 0, 
        main = "",
        main.cex = 2,
        sub.cex = 1,
        alpha = 0.50,
        # cat.pos=c(9,12)
    )
    grid.draw(venn.plot)
    # plot_grid(venn.plot, labels = c(method), ncol = 1, nrow = 1)
    
    TP = sum(pos_check & select_check)
    TN = sum(!pos_check & !select_check)
    FP = sum(!pos_check & select_check)
    FN = sum(pos_check & !select_check)
    
    recall = TP / (TP + FN)
    precision = TP / (TP + FP)
    f1 = 2 * precision * recall / (precision + recall) 
        
    results = data.frame(
        methods = c(method, method, method),
        val_type = c("precision", "recall", "f1"),
        val = c(precision, recall, f1)
    )
    
    for (each in list.files('./')){
      if (substring(each, nchar(each)-2) == 'log'){
        file.remove(each)
      }
    }
    return(results)
}
```

# analysis for selected cells of pencil
```{r}
pos_check = pbmc@meta.data[["true_groups_simulation"]]!='Group1'
select_check = pbmc@meta.data[["predicted_lables"]] != 'Rejected'

ggplot()
pencil_results = overlap_analysis(pbmc, pos_check, select_check, method = 'Pencil')
ggsave('./results/vennplot_for_pencil.pdf')
```


# DAseq
```{r}
labels_res <- c("class_1")
labels_nonres <- c("class_2")

# embeddings = pbmc@reductions[["umap"]]@cell.embeddings
input_genes_id = 1:2000

selected_genes = VariableFeatures(pbmc)[input_genes_id]
exp = pbmc@assays$RNA@scale.data
exp_for_selected = exp[selected_genes, ]
exp_for_selected = t(as.matrix(exp_for_selected))

library(DAseq)
python2use <- "C:/Users/Rentao/anaconda3/python.exe"
GPU <- 0

ts = Sys.time()
da_cells <- getDAcells(
  X = exp_for_selected,
  cell.labels = pbmc@meta.data[["cell_phenotype_labels_simulation"]],
  labels.1 = labels_res,
  labels.2 = labels_nonres,
  k.vector = seq(50, 500, 50),
  plot.embedding = pbmc@reductions[["umap"]]@cell.embeddings
)
te = Sys.time()
print(te - ts)


da_cells$pred.plot
da_cells$rand.plot
da_cells <- updateDAcells(
  X = da_cells, pred.thres = c(-0.2,0.2),
  plot.embedding = pbmc@reductions[["umap"]]@cell.embeddings
)
da_cells$da.cells.plot

da_selected_cells = c(da_cells[["da.up"]], da_cells[["da.down"]])
da_predict = rep('da_rejcted', ncol(pbmc))
da_predict[da_cells[["da.up"]]] = 'up'
da_predict[da_cells[["da.down"]]] = 'down'
pbmc <- AddMetaData(pbmc, metadata = da_predict, 'daseq.predict')
# pbmc$da_predict <- NULL
DimPlot(pbmc, group.by = 'daseq.predict', cols=c("gray","#3283FE","red"),order = c("up","down"))
ggsave('./results/da_predict_results.pdf')

write.csv(pbmc$daseq.predict, file = './results/daseq_predicted_labels.csv')
pos_check = pbmc@meta.data[["true_groups_simulation"]]!='Group1'
select_check = seq(1:ncol(pbmc)) %in% da_selected_cells

ggplot()
daseq_results = overlap_analysis(pbmc, pos_check, select_check, method = 'DAseq')
ggsave('./results/vennplot_for_daseq.pdf')
```

# Milo
```{r}
library(miloR)
library(SingleCellExperiment)
library(scater)
library(dplyr)
library(patchwork)
library(Seurat)
library(ggplot2)

split.samples <- function(sample, background_check='none'){
    # if 'none': random sample, else use background information.
    for (label in unique(sample)){
        if (background_check == 'none'){
            n = sum(sample==label)
            tmp = rep(sprintf('%s1', as.character(label)), n)
            tmp[sample.int(n, floor(n/2))] = sprintf('%s2', as.character(label))
            sample[sample==label] = tmp
        } else {
            tmp = sample[sample==label & background_check]
            n = sum(sample==label & background_check)
            tmp[1:n] = sprintf('%s1', as.character(label))
            tmp[sample.int(n, floor(n/2))] = sprintf('%s2', as.character(label))
            sample[sample==label & background_check] = tmp
            
            tmp_ = sample[sample==label & !background_check]
            n = sum(sample==label & !background_check)
            tmp_[1:n] = sprintf('%s1', as.character(label))
            tmp_[sample.int(n, floor(n/2))] = sprintf('%s2', as.character(label))
            sample[sample==label & !background_check] = tmp_
        }
    }

    return(sample)
    
}
    
input_genes_id = 1:2000
pbmc_small <- pbmc[VariableFeatures(pbmc), ]
pbmc_small <- CreateSeuratObject(counts = pbmc_small@assays[["RNA"]]@counts, meta.data = pbmc_small@meta.data)
# pbmc_samll <- NormalizeData(object = pbmc_small, normalization.method = "LogNormalize", scale.factor = 10000)
pbmc_small <- ScaleData(object=pbmc_small)
pbmc_small@reductions[['umap']] <- pbmc@reductions[['umap']]

ts = Sys.time()
pbmc_small <- RunPCA(object=pbmc_small, features=VariableFeatures(pbmc)[input_genes_id])

sample = pbmc@meta.data[["cell_phenotype_labels_simulation"]]
pbmc_small <- AddMetaData(pbmc_small, metadata = sample, col.name = 'condition')

background_check = pbmc@meta.data[["true_groups_simulation"]]=='Group1'
# sample <- split.samples(sample, background_check = 'none')
sample <- split.samples(sample, background_check = background_check)

pbmc_small <- AddMetaData(pbmc_small, metadata = sample, col.name = 'sample')


pbmc_small_sce <- as.SingleCellExperiment(pbmc_small)
pbmc_small_milo <- Milo(pbmc_small_sce)

pbmc_small_milo <- buildGraph(pbmc_small_milo, k = 30, d=30, reduced.dim = "PCA")
pbmc_small_milo <- makeNhoods(pbmc_small_milo, prop = 0.1, k = 30, d=30, refined = TRUE, reduced_dims = "PCA")
plotNhoodSizeHist(pbmc_small_milo)
pbmc_small_milo <- countCells(pbmc_small_milo, meta.data = data.frame(colData(pbmc_small_milo)), sample='sample')

head(nhoodCounts(pbmc_small_milo))
design_mat = data.frame(colData(pbmc_small_milo))[,c("sample", "condition")]
design_mat = distinct(design_mat)
row.names(design_mat) = design_mat$sample


print(design_mat)

pbmc_small_milo <- calcNhoodDistance(pbmc_small_milo, d=30)
da_results <- testNhoods(pbmc_small_milo, design = ~condition, design.df = design_mat, norm.method = c("logMS"), reduced.dim = "PCA")

da_results %>%
  arrange(- SpatialFDR) %>%
  head()

# pbmc_small_milo <- buildNhoodGraph(pbmc_small_milo)
# plotUMAP(pbmc_small_milo) + plotNhoodGraphDA(pbmc_small_milo, da_results, alpha=0.05) +
#   plot_layout(guides="collect")

te = Sys.time()
print(te - ts)

cell_ids = colnames(pbmc)
milo_selected = 1:ncol(pbmc_small)
nhoods_selected_pos = pbmc_small_milo@nhoods[, da_results$SpatialFDR < 0.05 & da_results$logFC > 0]
# logFC_of_selected_nhoods = da_predict$logFC[da_results$SpatialFDR < 0.05]
if (dim(as.matrix(nhoods_selected_pos))[2] == 1){
select_check_pos = nhoods_selected_pos == 1
} else{
    if (sum(nhoods_selected_pos)==0){
    select_check_pos = rep(FALSE, ncol(pbmc_small))
    } else{
    select_check_pos = rowSums(nhoods_selected_pos) > 0
    }
}
milo_selected[select_check_pos] = 'positive'

nhoods_selected_neg = pbmc_small_milo@nhoods[, da_results$SpatialFDR < 0.05 & da_results$logFC < 0]
# logFC_of_selected_nhoods = da_predict$logFC[da_results$SpatialFDR < 0.05]
if (dim(as.matrix(nhoods_selected_neg))[2] == 1){
select_check_neg = nhoods_selected_neg == 1
} else{
    if (sum(nhoods_selected_neg)==0){
    select_check_neg = rep(FALSE, ncol(pbmc_small))
    } else{
    select_check_neg = rowSums(nhoods_selected_neg) > 0
    }
}
milo_selected[select_check_neg] = 'negative'

select_check = select_check_neg | select_check_pos
pos_check = pbmc@meta.data[["true_groups_simulation"]]!='Group1'
milo_selected[!select_check] = 'milo-rejected'

ggplot()
milo_results = overlap_analysis(pbmc, pos_check, select_check, method = 'Milo')
ggsave('./results/vennplot_for_milo.pdf')

pbmc <- AddMetaData(pbmc, metadata = milo_selected, col.name = 'milo.predict')
DimPlot(pbmc, group.by = 'milo.predict', label = F)

DimPlot(pbmc, group.by = 'milo.predict', cols=c("gray","#3283FE","red"),order = c("positive","negative"))
ggsave('./results/milo_result.pdf')

DimPlot(pbmc, group.by="cell_phenotype_labels_simulation", reduction='umap', cells = cell_ids[select_check])

# pbmc <- AddMetaData(pbmc, metadata = milo_selected, col.name = 'milo.predict')
write.csv(pbmc$milo.predict, file = './results/milo_predicted_labels.csv')

```

# Meld
```{r}
expr_id = '2s'

# expr_id = "pca10-umap_radom_rate_0.2_radom_rate_0.3_radom_rate_0.4_radom_rate_0.5_radom_rate_0.6_radom_rate_0.7"
prediction_result_file = sprintf('../../results/Feldman_T_cell_for_select_genes/meld/meld_results_%s.csv', expr_id)
predict=read.csv(prediction_result_file, stringsAsFactors=F)

pbmc <- AddMetaData(pbmc, metadata = predict$predict, col.name = 'meld.predict') 
write.csv(pbmc$meld.predict, file = './results/meld_predicted_labels.csv')

pos_check = pbmc@meta.data[["true_groups_simulation"]]!='Group1'
select_check = pbmc$meld.predict != 'meld-rejected'

DimPlot(pbmc, group.by = 'meld.predict', cols=c("gray","#3283FE","red"),order = c("class_2","class_1"))
ggsave('./results/meld_result.pdf')

ggplot()
meld_results = overlap_analysis(pbmc, pos_check, select_check, method = 'Meld')
ggsave('./results/vennplot_for_meld.pdf')


expr_id = '2s_in300'

# expr_id = "pca10-umap_radom_rate_0.2_radom_rate_0.3_radom_rate_0.4_radom_rate_0.5_radom_rate_0.6_radom_rate_0.7"
prediction_result_file = sprintf('../../results/Feldman_T_cell_for_select_genes/meld/meld_results_%s.csv', expr_id)
predict=read.csv(prediction_result_file, stringsAsFactors=F)

pbmc <- AddMetaData(pbmc, metadata = predict$predict, col.name = 'meld.predict_in300') 
write.csv(pbmc$meld.predict, file = './results/meld_predicted_labels_in300.csv')

pos_check = pbmc@meta.data[["true_groups_simulation"]]!='Group1'
select_check = pbmc$meld.predict != 'meld-rejected'

DimPlot(pbmc, group.by = 'meld.predict_in300', cols=c("gray","#3283FE","red"),order = c("class_2","class_1"))
ggsave('./results/meld_result_in300.pdf')

ggplot()
meld_results = overlap_analysis(pbmc, pos_check, select_check, method = 'Meld')
ggsave('./results/vennplot_for_meld_in300.pdf')

```

```{r}
save(pbmc, file='Feldman_T_cell_bi-classification.RData')
```


#summary
```{r}
precision_for_select_all = sum(pos_check) / ncol(pbmc)
f1_for_select_all = 2*(precision_for_select_all * 1.0) / (precision_for_select_all + 1.0)
select_all_results = data.frame(
        methods = c('Select_all', 'Select_all', 'Select_all'),
        val_type = c("precision", "recall", "f1"),
        val = c(precision_for_select_all, 1.0, f1_for_select_all)
    )

results = rbind(pencil_results, daseq_results, milo_results, meld_results, select_all_results)  
# results = results[results$val_type!='precision', ]
ggplot(data=results, mapping=aes(x=methods, y=val, fill = val_type))+geom_col(position = 'dodge', width = 0.5)+ylim(0, 1)
ggsave('./results/method_comparison_with_all.pdf')

results = rbind(pencil_results, daseq_results, milo_results, meld_results)  
# results = results[results$val_type!='precision', ]
ggplot(data=results, mapping=aes(x=methods, y=val, fill = val_type))+geom_col(position = 'dodge', width = 0.5)+ylim(0, 1)
ggsave('./results/method_comparison.pdf')
```

#deg-umap
```{r}
PENCIL_R_vs_NR_DEGs_005 = FindMarkers(pbmc, ident.1="Responder", ident.2="Non-responder",group.by = "pred_labels",features = features_for_calculation)
Paper_G_vs_Bd_DEGs_005 = FindMarkers(pbmc, ident.1="CD8_G", ident.2="CD8_B",group.by = "paper.good.bad",logfc.threshold=log2FC_cutoff,features = features_for_calculation)
Patient_R_vs_NR_DEGs_005 = FindMarkers(pbmc, ident.1="Responder", ident.2="Non-responder",group.by = "Conditions",logfc.threshold=log2FC_cutoff,features = features_for_calculation)
rownames(PENCIL_R_vs_NR_DEGs_005) -> PENCIL_R_vs_NR_DEGs_005$gene
rownames(Paper_G_vs_Bd_DEGs_005) -> Paper_G_vs_Bd_DEGs_005$gene
rownames(Patient_R_vs_NR_DEGs_005) -> Patient_R_vs_NR_DEGs_005$gene
```

